Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election

### 2024-03-12 paper/us_election_2020.qmd
minor changes

In this paper, a logistic regression model is used to forecast the outcome of the 2020 election, with data from the Cooperative Election Study. Logistic regression is a great choice since it is used to predict binary outcomes, such as election results (Trump or Biden). Our analysis focuses on estimating the likelihood of victory for either Trump or Biden, based on various demographic, geographic, and socioeconomic factors captured in the CES data set.We selected data features:race, region that they live, employment status, and who they voted for. The estimand in this paper is the number of people who support Trump and Biden in reality. However, it is difficult to measure the exact number of people who support Trump and Biden since there are millions of people in the United States and not all of them with the rights to vote would actually vote. Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is trained using sample dataset from the 2020 election dataset from CES.

@@

# raw_ces2020[1:5, 1:5] |>
#   kable()
```

@@

### Model Justification

@@

The trace plot in @fig-mcmc-1 does not suggest anything out of the ordinary. Also, with the Rhat plot (@fig-mcmc-2), we can observe that everything is close to 1, and no more than 1.05.

@@

## Diagnostics

>>>>

In this paper, a logistic regression model is used to forecast the outcome of the 2020 election, with data from the Cooperative Election Study. Logistic regression is a great choice since it is used to predict binary outcomes, such as election results (Trump or Biden). Our analysis focuses on estimating the likelihood of victory for either Trump or Biden, based on various demographic, geographic, and socioeconomic factors captured in the CES data set.We selected data features: race, region that they live, employment status, and who they voted for. The estimand in this paper is the number of people who support Trump and Biden in reality. However, it is difficult to measure the exact number of people who support Trump and Biden since there are millions of people in the United States and not all of them with the rights to vote would actually vote. Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is trained using sample dataset from the 2020 election dataset from CES.

@@

raw_ces2020[1:5, 1:5] |>
  kable()
```

@@

## Model Justification

@@

The trace plot in @fig-trace does not suggest anything out of the ordinary. Also, with the Rhat plot (@fig-rhat), we can observe that everything is close to 1, and no more than 1.05.

@@

## Markov chain Monte Carlo Convergence Check
